Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks

### 2016-09-09 gm_pce/review_response.Rmd

### **Referee #1**

\
*Reviewer: Campany et al have presented a work on spatial distribution of leaf gas exchange variables and parameters as a function of light regime and how they reflect an optimal distribution of resources in the canopy. The most remarkable result of this study is the occurrence of a higher gs in shaded leaves than in sunny ones.  It is not clear to me the advantages of shaded leaves in having higher gs than ones at sun. I can see the point of enhancing An when sunflecks, but if the whole day is integrated it would result in a dramatic reduction of WUE, since shaded leaves with stomata open will transpire more or similarly to sunny ones (assuming that sunny leaves have higher temperature and therefore higher leaf-to-air VPD. Honestly, I cannot see any optimisation for canopy C in here, if this would be the case. As authors discussed, perhaps we should considered a dynamic light environment to understand the advantages of this strategy, but still I find difficult to conciliate with an optimal distribution of resources, in most cases very limiting. Data shows that there was an increase of gs, and hence, water use, without an enhancement of KL, and the relationship between gsmax and Vcmax, which usually found in canopies, but in this study is broken was broken.*

\
*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~L~ calculations, we have re-examined the calculation and subsequent analyses in the manuscript. Initially, *K*~L~ was calculated with transpiration rates in shade leaves under low light conditions. As shade leave were allowed to stabilize under high light conditions before they were detached and measured for water potential, the leaf water status reflected these conditions and not those of low light. We have corrected the calculation of *K*~L~ and made changes in the results and discussion sections. We have also clarified the calculation in the Methods section.

After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense in relation to the maintain hydraulic supply to the observed stomatal behavior. We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments. These corrected values also allow us to address the possible coordination of *K*~L~ and *g*~m~ reported by Flexas et al. 2013.
